Co-citations in context

### 2019-05-28 cocit_r.tex
Revised Introduction

\citep{uzzi_atypical_2013} used a novel approach for co-citation analysis of 17.9 million articles and their cited references from the Web of Science (WoS)  to characterize a subset of highly cited articles with respect to both novel and conventional combinations of prior research. Observed co-citation frequencies, mapped to journals, were computed across the dataset and normalized (shifted and scaled) to expected values generated by Monte Carlo simulations under a random graph model. These normalized journal pair frequencies were termed \emph{z-scores} (Materials and Methods). Thus, every article was associated with multiple z-scores corresponding with the journal pairs represented by its citations. For each article, positional statistics of these z-scores were calculated to describe conventionality: high conventionality (HC) if the median z-score for an article was greater than the median of median z-scores of all articles and low conventionality for the converse (LC). Similarly,  an article was deemed to have high novelty (HN) if the tenth percentile of its z-scores  was negative and low novelty (LN) for the converse. Accordingly, each article was labeled with respect to conventionality and novelty, e.g, HCHN (denoting that the article exhibits high conventionality and high novelty), with all four combinations being possible. 
\citep{uzzi_atypical_2013} observed that HCHN articles were twice as likely to be highly cited, suggesting that novel combinations of ideas flavoring a body of conventional thought were a recipe for impact across the scientific literature.

Key to Uzzi {\em et al.}, however, is their random graph model and its underlying assumptions. A careful examination of the model (as defined by the Monte Carlo simulation) reveals that random substitutions of references used to generate expected values are equiprobable, which means that a reference in an article can be substituted by any other reference randomly chosen from those published in the same year, irrespective of subject matter, disciplinary focus and citation count. 
For example, the model permits substitution of a reference in quantum physics with equal probability by a reference in quantum biology (i.e., the same field) or in some entirely different field, such as classical literature, evolutionary biology, or anthropology. Such substitutions poorly model documented citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}. In addition, under this random model, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once, which is inconsistent with the actual frequency of citation and also with the frequency distribution of citations characterized as power law or lognormal~\citep{stringer_statistical_2010,perline_strong_2005}.  
Lastly,  as we will show here, while the citation switching algorithm used in Uzzi  {\em et al.} preserves the number of publications and the number of references in each publication, it does not preserve the disciplinary proportions of cited references within the network.
 Accordingly, the expected value calculations generated by the \cite{uzzi_atypical_2013} simulations and used in characterizing journal pairs in terms of conventionality and novelty  can be reasonably questioned on grounds of model mis-specification.

@@

Despite different methods used to generate expected values, both \citep{uzzi_atypical_2013} and \citep{boyack_vs_uzzi_2014} measured observed co-citation frequencies across the scientific literature without disciplinary constraints and subsequently used normalized frequencies to examine disciplinary subsets. In contrast, we chose instead to first construct semantically related sets of documents (disciplinary networks) and to measure observed and expected frequencies within these networks, consistent with researchers' patterns of citing predominantly work from their own field.

Accordingly, we used keyword searches of the scientific literature to create three citation networks themed around broad search terms. Within these disciplinary frameworks, we calculated expected values using a random graph model that accounted for existing citation frequencies and an efficient Monte Carlo simulation algorithm that permitted a substantially greater number of simulations. We hypothesized that our approach would  reduce model misspecification and better simulate citation practice, in turn leading to different and more relevant conclusions. Our study on these semantically-themed citation networks reveals significantly different patterns of conventionality and novelty between  citation networks and disciplines that challenges the conclusion that HCHN articles have a greater probability of being represented in highly cited publications. 
Instead, we conclude that conventional thought as commonly defined in our study, 
in \citep{uzzi_atypical_2013}, and  \citep{boyack_vs_uzzi_2014}, is more likely to drive higher citations.
 \emph{these are just filler conclusions that must be carefully rewritten and added to- need Jim's conclusions to be stated here.}.

@@

 is the standard deviation of the frequencies generated by simulation. As a result of these calculations, each publication becomes associated with a set of z-scores corresponding to the journal pairs derived from pairwise combinations of its cited references and positional statistics (quantiles) of z-scores were calculated for each publication. Publications were also labeled according to conventionality and novelty. (i) HC if the median z-score exceeded the median of median z-scores for all publications. LC if the median z-score was equal to or less than the median of median z-scores for all publications (ii) HN if the tenth percentile of z-scores for a publication was less than zero. LN if the tenth percentile of z-scores for a publication was greater than zero. (Do we need a comment here that while we are using the term z-score, we do not imply that these are related to a normal distribution. Rather, we are using the terminology from Uzzi et al.?)

We investigated multiple definitions of hit articles, as did Uzzi et al., with hits defined as the 1\%, 2\%, 5\%, and 10\% top-cited articles.  Do we want to report on the percentile being used for novelty at both the 1 and 10th percentile, or just focus on the 10th percentile?

>>>>

\citep{uzzi_atypical_2013} used a novel approach for co-citation analysis of 17.9 million articles and their cited references from the Web of Science (WoS)  to characterize a subset of highly cited articles with respect to both novel and conventional combinations of prior research. Observed co-citation frequencies, mapped to journals, were computed across the dataset by Monte Carlo simulation under a random graph model and normalized (shifted and scaled) to expected values. These normalized journal pair frequencies were termed \emph{z-scores} (Materials and Methods). Thus, every article was associated with multiple z-scores corresponding with the journal pairs represented in its citations. For each article, positional statistics of these z-scores were calculated to describe conventionality: high conventionality (HC) if the median z-score for an article was greater than the median of median z-scores of all articles and low conventionality for the converse (LC). Similarly,  an article was deemed to have high novelty (HN) if the tenth percentile of its z-scores  was negative and low novelty (LN) for the converse. Accordingly, each article was labeled with respect to conventionality and novelty, e.g, HCHN (denoting that the article exhibits high conventionality and high novelty), with all four combinations being possible. 
\citep{uzzi_atypical_2013} observed that HCHN articles were twice as likely to be highly cited, suggesting that novel combinations of ideas flavoring a body of conventional thought were a recipe for impact across the scientific literature.

Key to Uzzi {\em et al.}, however, is their random graph model and its underlying assumptions. A careful examination of their Monte Carlo simulation reveals that random substitutions of references used to generate expected journal pair frequencies are equiprobable, which means that a reference in an article can be substituted by any other reference randomly chosen from those published in the same year, irrespective of subject matter, disciplinary focus and citation count. 
For example, the model permits substitution of a reference in quantum physics with equal probability by a reference in quantum biology (i.e., the same field) or in some entirely different field, such as classical literature, evolutionary biology, or anthropology. Such substitutions poorly model documented citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}. In addition, under this random model, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once, which is inconsistent with the actual frequency of citation and also with the power law or lognormal citation distributions observed in the literature~\citep{stringer_statistical_2010,perline_strong_2005}.  
Lastly,  as we will show here, while the citation switching algorithm used in Uzzi  {\em et al.} preserves the number of publications and the number of references in each publication, it does not preserve the proportion of cited references from within the disciplinary network.
 Accordingly, the expected value calculations generated by the \cite{uzzi_atypical_2013} simulations and used in characterizing journal pairs in terms of conventionality and novelty  can be reasonably questioned on grounds of model mis-specification.

@@

Despite different methods used to generate expected values, both \citep{uzzi_atypical_2013} and \citep{boyack_vs_uzzi_2014} measured observed co-citation frequencies across the scientific literature without disciplinary constraints and subsequently used normalized frequencies to examine disciplinary subsets. In contrast, we chose to first construct semantically related sets of documents (disciplinary networks) and to measure observed and expected frequencies within these networks in order to preserve the proportion of citations from a citing document's subject area under random sampling.

Accordingly, we used keyword searches of the scientific literature to create three citation networks themed around broad search terms. Within these disciplinary frameworks, we calculated expected journal pair frequencies using a random graph model that accounted for existing citation frequencies and an efficient Monte Carlo simulation algorithm that permitted a substantially greater number of simulations. We hypothesized that our approach would  reduce model misspecification and better simulate citation practice, in turn leading to different and more relevant conclusions. Our study on these semantically-themed citation networks reveals significantly different patterns of conventionality and novelty between  citation networks and disciplines that challenges the conclusion that HCHN articles have a greater probability of being represented in highly cited publications. 
Instead, we conclude that conventional thought as commonly defined in our study, 
in \citep{uzzi_atypical_2013}, and  \citep{boyack_vs_uzzi_2014}, is more likely to drive higher citations.
 \emph{these are just filler conclusions that must be carefully rewritten and added to- need Jim's conclusions to be stated here.}.

@@

 is the standard deviation of the frequencies generated by simulation. As a result of these calculations, each publication becomes associated with a set of z-scores corresponding to the journal pairs derived from pairwise combinations of its cited references and positional statistics (quantiles) of z-scores were calculated for each publication. Publications were subsequently labeled according to conventionality and novelty: (i) HC if the median z-score exceeded the median of median z-scores for all publications and LC if the median z-score was equal to or less than the median of median z-scores for all publications, 
and (ii) HN if the tenth percentile of z-scores for a publication was less than zero, and LN if the tenth percentile of z-scores for a publication was greater than zero. (Should we comment here that while we are using the term z-score, reusing it from Uzzi, we do not imply that these are related to a normal distribution?)

We investigated multiple definitions of hit articles, as did Uzzi et al., with hits defined as the 1\%, 2\%, 5\%, and 10\% top-cited articles.  Also, mention different thresholds, 1st and 10th, for novelty threshold.
